---
output:
  pdf_document: default
  html_document: default
  word_document: default
always_allow_html: true
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=FALSE}
setwd(PATH)
library(knitr)

Rmarkdown <- paste(PATH, "/R_scripts/ReportRMarkdown.R", sep = "")
read_chunk(Rmarkdown)
knitr::opts_chunk$set(echo = TRUE) 
```

```{r, cache= FALSE, results="hide", include=FALSE}
<<NAAM>>
```

```{r, echo=FALSE}
setwd(PATH)
library(knitr)

Configuration <- "Configuration.R"
Configuration <- file.path(path_data_in, Configuration)
read_chunk(Configuration)
knitr::opts_chunk$set(echo = TRUE) 
```

```{r, cache= FALSE, results="hide", include=FALSE}
<<statistics>>
 <<settings>>
```

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r out.width = "100%", fig.align="center"}
  setwd(PATH)
  library(EBImage)

  logo <- "Logo.png"
  logo <- file.path(PATH, 'Data/Reports/Template/Logo.png')
  LOGO <- readImage(logo)
  display(LOGO)
```

```{r out.width = "100%", fig.align="center"}
  setwd(PATH)
  library(EBImage)

  Title <- "PART2.png"
  Title <- file.path(PATH, 'Data/Reports/Template/PART2.png')
  TITLE <- readImage(Title)
  display(TITLE)

```

```{r}
Date <- Sys.Date()

cat(paste("This is the automatic report for the statistical analysis of experiment ", name_project, sep = ""))
cat("\n\n")
cat(paste("This report is made on ", Date, sep = ""))
```

\newpage

**Materials and methods**

The polarity that was used in this experiment is:
```{r, cache= FALSE, echo=FALSE}
cat(POLARITY)
```

The normalisation method that was performed is:
```{r, cache= FALSE, echo=FALSE}
cat(NORMALIZE_METHOD)
```

Retain all metabolites or remove specific metabolites before analysis: 
```{r, cache= FALSE, echo=FALSE}
cat(COEFICIENT_OF_VARIANCE)
cat(QC_PREFILTER)
```

The cut-off value for correlation in S-plot is:
```{r, cache= FALSE, echo=FALSE}
cat(CUTOFF_CORRELATION)
```

To check the normality Skewkurt plots were created.

Skewness-kurtosis plots indicate the range of skewness and kurtosis values a distribution can fit. Skewness is a measure of symmetry, or more precisely, the lack of symmetry and kurtosis is a measure of whether the data are heavy-tailed or light-tailed relative to a normal distribution. (Ho et al.2015)


The following methods were performed: 
```{r, cache= FALSE, echo=FALSE}

if(AMOUNT_OF_PROJECTIONS >= 1){
  cat("Principal component analysis (PCA)")
  cat("\n")
  cat("PCA reduces the dimensionality of a data set while retaining as much as possible of the variation.")
  cat("\n")
  cat("(Westerhuis et al., 2010)")
  cat("\n\n")
}
if(AMOUNT_OF_COMPARISONS >= 1){
  cat("Orthogonal partial least squares discriminant analysis (OPLS-DA)")
  cat("\n")
  cat("OPLS-DA is suitable for diagnosing differences between two groups or systems.")
  cat("\n")
  cat("It will tell us which variables have the largest discriminatory power.")
  cat("\n")
  cat("(Nisbet et al. 2018)")
  cat("\n\n") 
  
  cat("S plot")
  cat("\n")
  cat("This plot enables the extraction of statistically significant variables.")
  cat("\n")
  cat("It also enables the identification of shared/unique structures in the samples")
  cat("\n")
  cat("(Wiklund et al. 2008)")
  cat("\n\n")  
  
  cat("Variable Importance in Projection (VIP) plot")
  cat("\n")
  cat("VIP plot visualizes the importance of each variable and is often used for variable selection.")
  cat("\n")
  cat("A variable with a VIP Score close to or greater than 1 can be considered important in given model.")
  cat("\n")
  cat("(Tugizimana  et al. 2016)")
  cat("\n\n")
  
  cat("Loading plot with CI using bootstrap")
  cat("\n")
  cat("This plot shows the loadings with their confidence intervals.")
  cat("\n")
  cat("Bootstrap is used to calculate empirical CI's, regardless of the distribution (e.g. non-Gaussian).")
  cat("\n")
  cat("(K. 2012)")
  cat("\n\n")
  
  cat("Heatmap")
  cat("\n")
  cat("A heatmap visualizes the relative abundance of ions detected in each sample with color intensity")
  cat("\n")
  cat("Data from the collums represent the samples and data from the rows represent the metabolites.")
  cat("\n")
  cat("Rows and collumns are reordered, making it possible to view patterns among groups of samples")
  cat("\n")
  cat("(Grace et al. 2016)")
  cat("\n\n") 
}
if(AMOUNT_OF_MULTIPLE_COMPARISONS >= 1){
  cat("Partial Least Squares Discriminant Analysis (PLS-DA)")
  cat("\n")
  cat("PLS-DA is performed in order to sharpen the separation between groups of observations")
  cat("\n")
  cat("This method helps to understand which variables carry the class separating information.")
  cat("\n")
  cat("(Westerhuis et al. 2010)")
  cat("\n\n") 
  
  cat("Linear Models for microarray data (Limma)")
  cat("\n")
  cat("Limma is a package for differential expression analysis of data arising from microarray experiments")
  cat("\n")
  cat("(Smyth, 2005)")
  cat("\n\n") 
}

```

**Sample information**

```{r echo = FALSE}

setwd(PATH)

INPUT <- INPUT_SAMPLES
INPUT <- file.path(path_data_in, INPUT)
sampleMetadata <- read.table(INPUT, header=TRUE, sep="\t")

COLLUMN_NR_SAMPLENAMES <- 1
COLLUMN_NR_LAST <- COLLUMN_NR_LAST_BEFORE_COMPARISONS + AMOUNT_OF_COMPARISONS + AMOUNT_OF_MULTIPLE_COMPARISONS 
sampleMetadata[ , c(COLLUMN_NR_SAMPLENAMES:COLLUMN_NR_LAST)]
```

\newpage

**Results from multivariate data-analysis**


SKEWKURTPLOT BEFORE NORMALISATION-TRANSFORMATION-SCALING:
```{r out.width = "65%", fig.align="center"} 

setwd(PATH)
plot <- paste(name_project, "_skewkurtplot_samples.png", sep="")
plot <- file.path(path_data_out, plot)
knitr::include_graphics(plot)

```

SKEWKURTPLOT AFTER NORMALISATION-TRANSFORMATION-SCALING:
```{r comment=NA, out.width = "65%", fig.align="center"} 

setwd(PATH)
fig3 <- paste(name_project, "_skewkurtplot_samples_scaled_and_log-transformed.png", sep = "")
fig3 <- file.path(path_data_out, fig3)
knitr::include_graphics(fig3)

```

AMOUNT OF VARIABLES WITH NORMAL DISTRIBUTION:
```{r,  fig.cap = "Normalities report samples"} 

setwd(PATH)

NormalitiesSample <- paste(name_project, "_Report_normalities.txt", sep="")
NormalitiesSample <- file.path(path_data_out, NormalitiesSample)
outputSample <- read.table(NormalitiesSample, sep="\t", skip = 2, nrows = 1,  fill = TRUE, header= FALSE)

library(knitr)
library(kableExtra)
kable_input <- kable(outputSample)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))
```

```{r, fig.cap= "Normalities report samples after normalisation and transformation"}

setwd(PATH)

Normalities_after <- paste(name_project, "_Report_normalities.txt", sep = "")
Normalities_after <- file.path(path_data_out, Normalities_after)
outputAfter <- read.table(Normalities_after, sep="\t", skip=8, nrows = 1,  fill = TRUE)

library(knitr)
library(kableExtra)
kable_input <- kable(outputAfter)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))
```

\newpage

```{r comment=NA, echo=FALSE, warning=FALSE, out.width= '60%', fig.align= "center"}

AMOUNT_OF_PCA <- AMOUNT_OF_PROJECTIONS      #amount of pairwise comparisons, =0 if none

if(AMOUNT_OF_PCA >= 1){   
  cat("PCA SCOREPLOTS")
  cat("\n\n") 
  
  for(PCA in 1:AMOUNT_OF_PCA){
  cat(paste("PCA scoreplot ", PCA, sep="")) 
  setwd(PATH)
  library(EBImage)
  
  try({
  PCAscore <-  paste(name_project, "_PCA_scoreplot " , PCA, ".png", sep="")
  PCAscore <- file.path(path_data_out, PCAscore)
  PCAplot <- readImage(PCAscore)
  display(PCAplot)
  })
  }
}
```

\pagebreak

```{r comment=NA, echo=FALSE, warning=FALSE, out.width= '70%'}

if(AMOUNT_OF_COMPARISONS >= 1){
  cat("ORTHOGONAL PARTIAL LEAST SQUARES DISCRIMINANT ANALYSIS (OPLS-DA)")
  cat("\n\n")
  cat("If a model could be created, discriminating metabolites were determined.")
  cat("\n")
  cat("If no OPLS-DA model could be created, this is depicted by the comment 'object is NULL'.")
  cat("\n")
  cat("In this case dimensionreduction has led to one dimension, therefore no plot could be made.")
  
  for(pairwise_comparison in 1:AMOUNT_OF_COMPARISONS){
  
  setwd(PATH)
  library(EBImage)
  
  cat("\n\n")
  cat("******************************************************************************")
  cat("\n")
  cat(paste("OPLS-DA COMPARISON "), pairwise_comparison, sep = "")
  cat("\n")
  cat("******************************************************************************")
  cat("\n\n")
  
  try({
  name_plot <- paste(name_project, "_OPLSDA pairwise comparison " , pairwise_comparison, "_bis.png", sep="")
  name_plot <- file.path(path_data_out, name_plot)
  x <- readImage(name_plot)
  display(x)
  })
 
  cat("\n\n")
  cat(paste("OPLSDA scores comparison", pairwise_comparison, sep=""))
  cat("\n\n")
  
  try({
  OPLS <-  paste(name_project, "_Report_OPLSDA_scores_comparison_" , pairwise_comparison, ".txt", sep="")
  OPLS <- file.path(path_data_out, OPLS)
  OPLSDA <- read.table(OPLS, header = TRUE)
  print(OPLSDA)
  })
  
  cat("\n\n")
  cat(paste("VIP plot ", pairwise_comparison, sep = ""))
  
  try({
  name_VIP <-  paste(name_project, "_OPLSDA VIP-plot Comparison " , pairwise_comparison, ".png", sep="")
  name_VIP <- file.path(path_data_out, name_VIP)
  VIP <- readImage(name_VIP)
  display(VIP)
  })
  
  cat("\n\n")
  cat(paste("S plot ", pairwise_comparison, sep = ""))
  
  try({
  name_Splot <-  paste(name_project, "_OPLSDA S-plot Comparison " , pairwise_comparison, ".png", sep="")
  name_Splot <- file.path(path_data_out, name_Splot)
  S <- readImage(name_Splot)
  display(S)
  })
  
  cat("\n\n")
  cat(paste("Loading plot ", pairwise_comparison, sep= ""))
  
  try({
  name_loadingplot <-  paste(name_project, "_OPLSDA loadingplot Comparison " , pairwise_comparison, ".png", sep="")
  name_loadingplot <- file.path(path_data_out, name_loadingplot)
  L <- readImage(name_loadingplot)
  display(L)
  })
  
  cat("\n\n")
  cat(paste("HEATMAP ", pairwise_comparison, sep= ""))
  
  try({
  name_heatmap <-  paste(name_project, "_OPLSDA Heatmap comparison ", pairwise_comparison, ".png", sep="")
  name_heatmap <- file.path(path_data_out, name_heatmap)
  H <- readImage(name_heatmap)
  display(H)
  })
  
  cat("\n\n")
  cat(paste("Table selected metabolites ", pairwise_comparison, sep= ""))
  cat("\n\n")
  
  try({
  setwd(PATH) 
  Selection <-  paste(name_project, "_Variables_after_selection_VIP1_CIno0_Corr_cutoff_comparison_" , pairwise_comparison, ".txt", sep="")
  Selection <- file.path(path_data_out, Selection)
  SEL <- read.table(Selection, header = TRUE, sep = "")
  print(SEL[ , c(1:4, 20,21,23, 25:27)])
  })
  }
}
```

\pagebreak

```{r comment=NA, echo=FALSE, warning=FALSE, out.width='70%'}

AMOUNT_OF_MULTICOMPARISONS <- AMOUNT_OF_MULTIPLE_COMPARISONS    #amount of pairwise comparisons, =0 if none

if(AMOUNT_OF_MULTICOMPARISONS >= 1){   

  for(multiple_comparison in 1:AMOUNT_OF_MULTICOMPARISONS){
  
  setwd(PATH)
  library(EBImage)
  cat("PARTIAL LEAST SQUARES DISCRIMINANT ANALYSIS (PLS-DA)")
  cat("\n\n")
  cat("If a model could be created, a scoreplot is presented.")
  cat("\n")
  cat("If no PLS-DA model could be created, this is depicted by the comment 'object is NULL'. ")
  cat("\n")
  cat("In this case dimensionreduction has led to one dimension, therefore no scoreplot could be made.")
  cat("\n\n")
  cat("******************************************************************************")
  cat("\n")
  cat(paste("PLSDA multiple comparison ", multiple_comparison, sep=""))
  cat("\n")
  cat("******************************************************************************") 
  
  try({
  name_multi <-  paste(name_project, "_PLSDA multiple comparison " , multiple_comparison, "_bis.png", sep="")
  name_multi <- file.path(path_data_out, name_multi)
  Multi <- readImage(name_multi)
  display(Multi)
  })

  try({
  cat("\n\n")
  cat(paste("PLS-DA scores for multiple comparison ", multiple_comparison, sep = "" ))
  cat("\n\n")
  multiscore <- paste(name_project, "_Report_PLSDA_scores_multiple_comparison_" , multiple_comparison, ".txt", sep="")
  multiscore <- file.path(path_data_out, multiscore)
  MULTI <- read.table(multiscore, header = TRUE)
  print(MULTI)
  }) 

  try({
  cat("\n\n")
  cat(paste("Toptable multiple comparison ", multiple_comparison, sep = "" ))
  cat("\n\n")
  cat("Analysis with Limma (Linear Models for Microarray Data)")
  cat("\n\n")
  TOPTABLE <- paste(name_project, "_ToptableF_output_multiple_comparison" , multiple_comparison, ".txt", sep="")
  TOPTABLE <- file.path(path_data_out, TOPTABLE)
  outputMultiTable <- read.table(TOPTABLE, header=TRUE, sep="\t")
  MultiTOP <- outputMultiTable[c(1:30), c(1:14)]
  print(MultiTOP)
  }) 
  }
}
```


